A Toy Model for Inverse Covariance Weighting

### 2018-10-09 Memo1_inversecov.tex
adding memos corresponding to 3 appendices

\documentclass[preprint2,numberedappendix,tighten]{aastex6}  % USE THIS TO MAKE BIB, THEN FORMAT USING EMULATEAPJ
%\documentclass[twocolumn,numberedappendix]{emulateapj}
\shorttitle{Characterizing Signal Loss in the 21\,cm Reionization Power Spectrum}
\shortauthors{Cheng et al.}

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[figuresright]{rotating}
\usepackage{natbib}
\usepackage{ctable}
\usepackage{comment}
\usepackage{bm}
%\usepackage{mathtools}
\usepackage{xparse}
%\usepackage{titlesec}
\citestyle{aa}

@@

%Math stuff from Danny
\NewDocumentCommand{\evalat}{sO{\big}mm}{%
  \IfBooleanTF{#1}
   {\mleft. #3 \mright|_{#4}}
   {#3#2|_{#4}}%
}

%		Math Shortcuts from Adrian
%\def\b{\mathbf{b}}
%\def\k{\mathbf{k}}
%\def\r{\mathbf{r}}
%\def\q{\mathbf{q}}
%\def\b{\mathbf{b}}
%\def\kp{\mathbf{k}^\prime}
%\def\kpp{\mathbf{k}^{\prime\prime}}
%\def\V{\mathbb{V}}
%\def\At{\tilde{A}}
%\def\Vt{\tilde{V}}
%\def\Tt{\tilde{T}}
%\def\tb{\langle T_b\rangle}
%\newcommand{\vis}{\mathbf{v}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\f}{\mathbf{f}}
\newcommand{\s}{\mathbf{s}}
\newcommand{\e}{\mathbf{e}}
\newcommand{\p}{\mathbf{p}}
\newcommand{\phat}{\widehat{\mathbf{p}}}
\newcommand{\C}{\mathbf{C}}
\newcommand{\Chat}{\mathbf{\widehat{C}}}
\newcommand{\F}{\mathbf{F}}
\newcommand{\Fhat}{\mathbf{\widehat{F}}}
\newcommand{\Q}{\mathbf{Q}}
\newcommand{\I}{\mathbf{I}}
%\newcommand{\xhat}{\widehat{\mathbf{x}}}
%\newcommand{\nhat}{\widehat{\mathbf{n}}}
%\newcommand{\A}{\mathbf{A}}
%\newcommand{\N}{\mathbf{N}}
%\newcommand{\rhat}{\widehat{\mathbf{r}}}
%\newcommand{\khat}{\widehat{\mathbf{k}}}
%\newcommand{\btheta}{\boldsymbol \theta}

\newcommand{\cc}[1]{{\color{purple} \textbf{[CC: #1]}}}
\newcommand{\acl}[1]{{\color{red} \textbf{[ACL:  #1]}}}
\newcommand{\dcj}[1]{{\color{orange} \textbf{[DCJ: #1]}}}
\newcommand{\mjk}[1]{ {\color{olive} \textbf{[MJK: #1]}} }
\newcommand{\jea}[1]{{\color{blue} \textbf{[JEA: #1]}} }
\newcommand{\jp}[1]{ {\color{pink} \textbf{[JCP: #1]}} } %y'all picked all the good colors

% Math shortcuts from James
\newcommand{\invC}{\ensuremath{\C^{-1}}}
\newcommand{\dC}[1]{\ensuremath{\C_{,#1}}}
\DeclareMathOperator{\Tr}{tr}
\newcommand{\half}{\ensuremath{\frac{1}{2}}}
\newcommand{\PDeriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}

% Math shortcuts from Danny
\newcommand{\Prob}{\mathbf{P}}

\begin{document}
\title{Characterizing Signal Loss in the 21\,cm Reionization Power Spectrum: \\A Revised Study of PAPER-64}

\author{
Carina Cheng\altaffilmark{1}$^{,\diamond}$,
Aaron R. Parsons\altaffilmark{1,2},
Matthew Kolopanis \altaffilmark{3},
Daniel C. Jacobs\altaffilmark{3}, 
Adrian Liu\altaffilmark{1,4}$^{,\dagger}$, 
Saul A. Kohn\altaffilmark{5},
James E.~Aguirre\altaffilmark{5},
Jonathan C. Pober\altaffilmark{6}, 
Zaki S. Ali\altaffilmark{1}, 
Gianni Bernardi\altaffilmark{7,8,9}, 
Richard F. Bradley\altaffilmark{10,11,12},
Chris L. Carilli\altaffilmark{13,14},
David R. DeBoer\altaffilmark{2}, 
Matthew R. Dexter\altaffilmark{2},
Joshua S. Dillon\altaffilmark{1}$^{,*}$,
Pat Klima\altaffilmark{11},
David H. E. MacMahon\altaffilmark{2},
David F. Moore\altaffilmark{5},
Chuneeta D. Nunhokee\altaffilmark{8},
William P. Walbrugh\altaffilmark{7},
Andre Walker\altaffilmark{7}
}

\altaffiltext{1}{Astronomy Dept., U. California, Berkeley, CA}
\altaffiltext{2}{Radio Astronomy Lab., U. California, Berkeley CA}
\altaffiltext{3}{School of Earth and Space Exploration, Arizona State U., Tempe AZ}
\altaffiltext{4}{Berkeley Center for Cosmological Physics, Berkeley, CA} 
\altaffiltext{5}{Dept. of Physics and Astronomy, U. Penn., Philadelphia PA} 
\altaffiltext{6}{Dept. of Physics, Brown University, Providence RI}
\altaffiltext{7}{Square Kilometer Array, S. Africa, Cape Town South Africa}
\altaffiltext{8}{Dept. of Physics and Electronics, Rhodes University, South Africa}
\altaffiltext{9}{INAF-Instituto di Radioastronomia, Bologna Italy}
\altaffiltext{10}{Dept. of Electrical and Computer Engineering, U. Virginia, Charlottesville VA}
\altaffiltext{11}{National Radio Astronomy Obs., Charlottesville VA}
\altaffiltext{12}{Dept. of Astronomy, U. Virginia, Charlottesville VA}
\altaffiltext{13}{National Radio Astronomy Obs., Socorro NM}
\altaffiltext{14}{Cavendish Lab., Cambridge UK}

%		Notes

%Reference section with: \ref{sec:Intro}
%Reference equation with: \eqref{eq:eqtest}
%Reference figure with: \ref{fig:figtest}
%Cite paper inside sentence: \citet{ref}
%Cite paper at end of sentence: \citep{ref}
%Cite paper inside a parenthetical sentence: \citealt{ref}

%To compile with references shown, compile in BibTeX once and LaTeX twice

%		Sample Equation Syntax
%\begin{equation}
%\label{eqtest}
%\langle \widetilde{T} (\mathbf{k}) \widetilde{T}^* (\mathbf{k^\prime}) \rangle = (2 \pi)^3 \delta^D (\mathbf{k} - \mathbf{k}^\prime) P(k),
%\end{equation}

% ABSTRACT ---------------------------------------------------------------------------------

\begin{abstract}
The Epoch of Reionization (EoR) is an uncharted era in our Universe's history during which the first stars and 
galaxies led to the ionization of neutral hydrogen in the intergalactic medium. There are many experiments investigating the 
EoR by tracing the 21\,cm line of neutral hydrogen. Because this signal is very faint and difficult to isolate, it is crucial to develop analysis techniques that maximize sensitivity and suppress contaminants in data. It is also imperative to understand the trade-offs between different analysis methods and their effects on power spectrum estimates. Specifically, with a statistical power spectrum detection in HERA's foreseeable future, it has become increasingly important to understand how certain analysis choices can lead to the loss of the EoR signal. In this paper, we focus on signal loss associated with power spectrum estimation. We describe the origin of this loss using both toy models and data taken by the 64-element configuration of the Donald C. Backer Precision Array for Probing the Epoch of Reionization (PAPER). In particular, we highlight how detailed investigations of signal loss have led to a revised, higher 21\,cm power spectrum upper limit from PAPER-64. Additionally, we summarize errors associated with power spectrum error estimation that were previously unaccounted for. We focus on a subset of PAPER-64 data in this paper; revised power spectrum limits from the PAPER experiment are presented in a forthcoming paper by Kolopanis et al. (\textit{in prep.}) and supersede results from previously published PAPER analyses.
\end{abstract}

% INTRO ---------------------------------------------------------------------------------

\section{Introduction}
\label{sec:Intro}

{\let\thefootnote\relax\footnote{$^{\diamond}$\href{mailto:ccheng@berkeley.edu}{ccheng@berkeley.edu}}}
{\let\thefootnote\relax\footnote{$^{\dagger}$Hubble Fellow}}
{\let\thefootnote\relax\footnote{$^{*}$NSF AAPF Fellow}}
\setcounter{footnote}{0}

By about one billion years after the Big Bang ($z \sim 6$), the first stars and galaxies are thought to have ionized all the 
neutral hydrogen that dominated the baryonic matter content in the Universe. This transition period, during which the first 
luminous structures formed from gravitational collapse and began to emit intense radiation that ionized the cold neutral gas 
into a plasma, is known as the Epoch of Reionization (EoR). The EoR is a relatively unexplored era in our Universe's history, which spans the birth of the first stars to the full reionization of the intergalactic medium (IGM). This epoch encodes important information regarding the nature of the first galaxies and the processes of structure formation. 
Direct measurements of the EoR would unlock powerful characteristics about the IGM, revealing connections 
between the matter distribution exhibited via cosmic microwave background (CMB) studies and the highly structured 
web of galaxies we observe today (for a review, see \citet{barkana_and_loeb2001}, \citet{furlanetto_et_al2006} and \citet{loeb_furlanetto_2013}).

One promising technique to probe the EoR is to target the 21\,cm wavelength signal that is emitted and absorbed by neutral hydrogen via 
its spin-flip transition (\citealt{furlanetto_et_al2006}; \citealt{barkana_and_loeb2008}; \citealt{morales_and_wyithe2010}; \citealt{pritchard_and_loeb2010}; \citealt{pritchard_loeb2012}). This technique is powerful because it can be observed both spatially and as a function of redshift --- that is, the wavelength 
of the signal reaching our telescopes can be directly mapped to a distance from where the emission originated before 
stretching out as it traveled through expanding space. Hence, 21\,cm tomography offers a unique window into both the spatial and temporal
evolution of ionization, temperature, and density fluctuations.

In addition to the first tentative detection of our \textit{Cosmic Dawn} (pre-reionization era) made by the Experiment to Detect the Global EoR Signature (EDGES; \citealt{bowman_et_al2018}; \citealt{bowman2010}), there are several radio telescope experiments that have succeeded in using 
the 21\,cm signal from hydrogen to place constraints on the brightness of the signal. Examples of experiments investigating the 
mean brightness temperature of the 21\,cm signal relative to the CMB are the Large Aperture Experiment to Detect the Dark Ages (LEDA; \citealt{bernardi_et_al2016}), the 
Dark Ages Radio Explorer (DARE; \citealt{burns2012}), the Sonda Cosmol\'ogica de las Islas para la Detecci\'on de 
Hidr\'ogeno NeutroSciHi (SCI-HI; \citealt{voytek2014}), the Broadband Instrument for Global HydrOgen ReioNisation Signal 
(BIGHORNS; \citealt{sokolowski2015}), and the Shaped Antenna measurement of the background RAdio Spectrum (SARAS; 
\citealt{patra2015}). Radio interferometers which seek to measure statistical power spectra include the Giant Metre-wave 
Radio Telescope (GMRT; \citealt{paciga_et_al2013}), the LOw Frequency ARray (LOFAR; \citealt{van_haarlem_et_al2013}), 
the Murchison Widefield Array (MWA; \citealt{tingay_et_al2013}), the 21 Centimeter Array (21CMA; 
\citealt{peterson_et_al2004}; \citealt{wu2009}), the Square Kilometre Array (SKA; \citealt{koopmans_et_al2015}), and PAPER (\citealt{parsons_et_al2010}). The Hydrogen Epoch of 
Reionization Array (HERA), which is currently being built, is a next-generation instrument that aims to combine lessons 
learned from previous experiments and has a forecasted sensitivity capable of a high-significance power spectrum 
detection with an eventual $350$ elements using current analysis techniques (\citealt{pober_et_al2014}; \citealt{liu_parsons_2016}; \citealt{dillon_parsons2016}; \citealt{deboer_et_al2017}).

The major challenge that faces all 21\,cm experiments is isolating a small signal that is buried underneath foregrounds and 
instrumental systematics that are, when combined, four to five orders of magnitude brighter \citep[e.g.,][]{santos_et_al2005, ali_et_al2008, deOliveiraCosta_et_al2008, jelic_et_al2008, bernardi_et_al2009, bernardi_et_al2010, ghosh_et_al2011, pober_et_al2013b, bernardi_et_al2013, dillon_et_al2014, kohn_et_al2016}. A clean measurement therefore requires an intimate understanding of how data analysis choices, which are often tailored to maximize sensitivity and minimize contaminants, affect power spectrum results. More specifically, it is imperative to develop techniques that ensure 
the accurate extraction and recovery of the EoR signal, despite the analysis method chosen and how much loss (of both contaminants and the EoR signal) accompanies the method. In this paper, we specifically discuss signal loss --- the loss of the \textit{cosmological} signal --- associated with power spectrum estimation. This is an issue that is essential to investigate for a robust 21\,cm power spectrum analysis and one that has motivated a revised PAPER analysis. We first approach this topic from a broad perspective, and then perform a detailed 
case study using data from the 64-element configuration of PAPER. In this study we use a subset of PAPER-64 data to illustrate our revised analysis methods, while a related paper, Kolopanis et al. (\textit{in prep.}), builds off of the methods in this paper to present revised PAPER-64 results for multiple redshifts and baseline types.

Finally, we also highlight several additional errors made in previous PAPER analyses, including those related to bootstrapping and error estimation. This paper accompanies the erratum of \citet{ali_et_al2018} and adds to the growing foundations of lessons which have been documented, for example, in \citet{Paciga2013}, \citet{Patil2016}, and \citet{Jacobs2016}, by the GMRT, LOFAR, and MWA projects respectively. These lessons are imperative as the community as a whole moves towards higher sensitivities and potential EoR detections.

This paper is organized into three main sections. In Section \ref{sec:SiglossOverview} we use toy models to develop intuition about signal loss, its origin, and its subtleties. In Section \ref{sec:CaseStudy}, we present a case study using data from the PAPER-64 array, highlighting key changes from the signal loss methods used in the published result in \citet{ali_et_al2015}, henceforth known as \citetalias{ali_et_al2015}, which previously underestimated loss. In Section \ref{sec:OtherErrors}, we summarize additional lessons learned since \citetalias{ali_et_al2015} that have shaped our revised analysis. Finally, we conclude in Section \ref{sec:Con}.

% SECTION 2 SIGLOSS TOY MODEL ---------------------------------------------------------------------------------
%\color{blue}

\section{Signal Loss Toy Models}
\label{sec:SiglossOverview}
% JEA - We never totally nail down what ``signal loss'' is ...

Signal loss refers to attenuation of the target cosmological signal 
in a power spectrum estimate. Certain analysis techniques can cause this loss, and if the amount of loss is not quantified accurately, it could lead to false non-detections and overly aggressive upper limits. Determining whether an analysis pipeline is lossy, and estimating the amount of loss if so, has subtle challenges but is necessary to ensure the accuracy of any result.

One type of signal loss can occur when weighting data by itself. Broadly speaking, a dataset can be weighted to emphasize certain features and minimize others. For example, one flavor of weighting employed by previous PAPER analyses is inverse covariance weighting in frequency, which is a generalized version of inverse variance weighting that also takes into account frequency correlations (\citealt{liu_tegmark2011}; \citealt{dillon_et_al2013a}; \citealt{liu_et_al2014a}; \citealt{liu_et_al2014b}; \citealt{dillon_et_al2014}; \citealt{dillon_et_al2015}). Using such a technique enables the down-weighting of contaminant modes that obey a different covariance structure from that of cosmological modes. However, a challenge of inverse covariance 
weighting is in estimating a covariance matrix that is closest to the true covariance of the data; the discrepancy between the two, as we will see, can have large impacts on signal loss. In this paper we focus specifically on loss associated with the use of an empirically estimated covariance matrix with the ``optimal quadratic estimator'' formalism.
This loss was significantly 
%applying a weighting matrix to data, a loss that can be significant depending on the choice of weighting and one that was previously 
underestimated in the \citetalias{ali_et_al2015} analysis and is the main reason motivating a revised power spectrum result.

% JEA - Put all of the (O)QE / ICW theory here, rather than sprinkled throughout this section.

\subsection{The Quadratic Estimator Method}
\label{sec:QE}

%AAAICW

We begin with an overview of the quadratic estimator (QE) formalism used for power spectrum estimation. The goal of power spectrum analysis is to produce an unbiased estimator of the EoR power spectrum in the presence of both noise and foreground emission. Prior to power spectrum estimation, the data will often have been prepared to have minimal foregrounds by some method of subtraction, so this foreground emission may appear either directly (because it was not subtracted) or as a residual of some subtraction process not in the power spectrum domain. If an accurate estimate of the total covariance of the data is known, including both the desired signal and any contaminants, then the ``optimal quadratic estimator'' formalism provides a method of producing a minimum variance, unbiased estimator of the desired signal, as shown in 
% JEA - I cut this b/c I wanted to step back and explain a bit more clearly what we were actually hoping for 
%Driven by the need to mitigate foreground bias, PAPER's previous analyses use a weighting method that aims to down-weight foregrounds. This weighting is applied to data, which is then propagated into a final estimator using the power spectrum estimation technique of quadratic estimators (QE) as done in 
\citet{liu_tegmark2011}, \citet{dillon_et_al2013a}, \citet{liu_et_al2014a}, \citet{liu_et_al2014b}, \citet{trott_et_al2012}, \citet{dillon_et_al2014}, \citet{dillon_et_al2015}, \citet{switzer_et_al2015}, and \citet{trott_et_al2016}. 
%Before showing how signal loss can arise when using different weighting matrices, we first summarize QE as performed in the PAPER analysis.

% JEA - The discussion of dimensions here is confusing ... "a baseline" is mentioned, but it's not clear if the redundant baseline averaging has already happened or what.  I believe we are only considering covariance between frequencies, but this isn't explicitly said, and the matrix equations don't make sense unless all the dimensions are N_f N_t.  
%We begin with our 
Suppose that the measured visibilities for a single baseline in Jy are arranged as a data vector, $\textbf{x}$. It has length $N_{t} N_{f}$,
% JEA - not sure this helps
%, but in practice we manipulate it as an array with dimensions ($N_{f}, N_{t}$), 
where $N_{t}$ is the number of time integrations and $N_{f}$ is the number of frequency channels. The covariance of the data is given by 
\begin{equation}
\textbf{C} \equiv \langle\textbf{xx}^{\dagger}\rangle = \textbf{S} + \textbf{U}
\end{equation}
where the average over an ensemble of data realizations produces the true covariance, and we further assume it may be written as the sum of the desired cosmological signal $\textbf{S}$ and other terms $\textbf{U}$.

We are interested in estimating the three-dimensional power spectrum of the EoR.  
Visibilities are measurements of the Fourier transform of the sky along two spatial dimensions (using the flat-sky approximation), and since we are interested in three-dimensional Fourier modes we only need to take one Fourier transform of our visibilities along the line-of-sight dimension.  We consider band powers $P^\alpha$ of the power spectrum of $\textbf{x}$ over some range in cosmological $\mathbf{k}$, where $\alpha$ indexes a waveband in $k_{\parallel}$ (a cosmological wavenumber $k_{\parallel}$ is the Fourier dual to frequency under the delay approximation (\citealt{parsons_et_al2012b}), which is a good approximation for the short baselines that PAPER analyzes).  The fundamental dependence of the covariance on the power spectrum band powers $P^\alpha$ is encoded as 
% JEA - Should distinguish slightly more clearly between Q and dC/dp_alpha
\begin{equation}
\textbf{S} = \sum_\alpha P^\alpha \frac{\partial\textbf{C}}{\partial P^\alpha} \equiv \sum_\alpha P^\alpha \textbf{Q}^\alpha
%\textbf{S} = p_{\alpha} \frac{\partial\textbf{C}}{\partial p_{\alpha}}
%\frac{\partial\textbf{C}}{\partial p_{\alpha}} \equiv \textbf{Q}^{\alpha}
\end{equation}
where we define $\frac{\partial\textbf{C}}{\partial P^\alpha} \equiv \textbf{Q}^{\alpha}$. In other words, $\textbf{Q}$ describes the response of the covariance to a change in the power spectrum, relating a quadratic statistic of the data (the covariance) to a quadratic statistic in Fourier-space (the power spectrum). 
% JEA - ah, the dimensions again
% which has dimensions ($N_{f}$,$N_{f}$).

The optimal quadratic estimator prescription is then to compute
\begin{equation}
\label{eq:OQE}
\widehat{P}^{\alpha}  = \sum_\beta ({\F^{-1}})^{\alpha\beta} (\widehat{q}^{\beta} - \widehat{b}^{\beta} )
\end{equation}
where $\F$ is the Fisher matrix (which determines errors on the power spectrum estimate)
\begin{equation}
F^{\alpha \beta} \equiv \frac{1}{2} \textrm{tr} \left( \C^{-1} \textbf{Q}^{\alpha} \C^{-1} \textbf{Q}^{\beta} \right),
\end{equation}
$\widehat{q}$ is the un-normalized power spectrum estimate
\begin{equation}
\label{eq:OQEQuadratic}
\widehat{q}^\alpha =  \half \textbf{x}^\dagger \invC \textbf{Q}^{\alpha}  \invC \textbf{x},
\end{equation}
and $\widehat{b}$ is the additive bias
\begin{equation}
\label{eq:OQELinear}
\widehat{b}^{\alpha} = \half \Tr\left( \mathbf{U} \invC \textbf{Q}^{\alpha} \invC \right).
\end{equation}
The power spectrum estimator in Equation \eqref{eq:OQE} is the minimum variance (smallest error bar) estimate of the power spectrum subject to the constraint that it is also unbiased; that is, the ensemble average of the estimator is equal to its true value
\begin{equation}
\label{eq:super_unbiased}
\langle \widehat{P}^{\alpha} \rangle = P^\alpha
\end{equation}
\citep{tegmark_et_al1997a,bond_et_al1998}.

Intuitively, the estimator must be capable of ``suppressing" or ``removing" the effects of contaminants in order to obtain an unbiased estimate of the power spectrum. By construction, the subtraction of the residual foreground and noise bias accomplishes this, removing any additive bias. However, the $\mathbf{C}^{-1}$ piece of Equation \eqref{eq:OQEQuadratic} also has the effect of suppressing residual foregrounds and noise, in both the additive bias and any contributions the residuals may have to the variance.

More specifically, the effect of the weighting in Equation \eqref{eq:OQEQuadratic} is to project out the modes of $\textbf{U}$ with a different covariance structure than $\mathbf{S}$ in the power spectrum estimate, and the effect of Equation \eqref{eq:OQELinear} is to subtract out the remaining bias. Similar effects for a realistic model of the EoR and foregrounds are shown in \citet{liu_tegmark2011}. If the covariance structure of the contaminants is sufficiently different from the desired power spectrum, then the linear bias term may be expected to be quite small, and it is only necessary to know $\C$ and $\textbf{Q}^{\alpha}$, but not $\mathbf{U}$. Since the foregrounds are expected to be strongly correlated between frequencies whereas the EoR is not, we expect different covariance structures and therefore a small linear bias. Moreover, because the linear bias is always positive and there is no multiplicative bias, the quadratic-only term will always produce an estimate which is {\it high} relative to the true value, and which can conservatively be interpreted as an upper limit. These considerations, and the difficulty of obtaining an estimate for $\mathbf{U}$, motivate the neglect of the linear bias in the rest of this analysis.

%We do this when forming the un-normalized power spectrum estimate $\widehat{q}_{\alpha}$:
Motivated by the desire to retain the advantageous behavior of suppressing contributions of $\mathbf{U}$ to estimates of the EoR power spectrum, we note that is possible to define a modified version of the quadratic estimator 
%which may still produce an unbiased estimate, 
where Equation \eqref{eq:OQEQuadratic} is replaced by
\begin{equation}
\label{eq:qhat}
\widehat{q}^{\alpha} = \frac{1}{2}\textbf{x}^{\dagger}\textbf{R}\textbf{Q}^{\alpha}\textbf{R}\textbf{x}
\end{equation}
where $\textbf{R}$ is a weighting matrix chosen by the data analyst.  For example, inverse covariance weighting (the optimal form of QE) would set $\textbf{R} \equiv \textbf{C}^{-1}$ and a uniform-weighted case would use $\textbf{R} \equiv \textbf{I}$, the identity matrix.
Again, the matrix $\textbf{Q}^{\alpha}$ encodes the dependence of the covariance on the power spectrum but in practice also also does other things, including implementing a transform of the frequency domain visibilities to $\mathbf{k}$-space, taking into account cosmological scalings, and converting the visibilities from Jansky to Kelvin.
%with dimensions ($N_{f}$,$N_{f}$) --- as

%%% COME BACK HERE
% Need to still explain the motivation for this
With an appropriate normalization matrix $\textbf{M}$, the quantity
%We normalize our power spectrum estimates using the matrix $\textbf{M}$:
\begin{equation}
\label{eq:phat}
\widehat{\textbf{P}} = \textbf{M}\widehat{\textbf{q}}
\end{equation}
is a sensible \textit{estimate} of the \textit{true} power spectrum $\textbf{P}$.

To ensure that $\textbf{M}$ correctly normalizes our power spectrum, one may take the expectation value of Equation \eqref{eq:phat} to obtain
\begin{eqnarray}
\langle \widehat{P}^\alpha \rangle &=& \frac{1}{2} \sum_{\beta \gamma} M^{\alpha \gamma}  \textrm{tr} \left( \textbf{R}\textbf{Q}^{\gamma}\textbf{R} \textbf{Q}^{\beta}\right) P^\beta +\frac{1}{2} \sum_{ \gamma}  \Tr\left( \mathbf{U} \textbf{R} \textbf{Q}^{\gamma} \mathbf{R} \right) \nonumber \\
&\equiv & \sum_{\beta} W^{\alpha \beta} P^\beta +\frac{1}{2} \sum_{ \gamma}  \Tr\left( \mathbf{U} \textbf{R} \textbf{Q}^{\gamma} \mathbf{R} \right), \label{eq:Wpplusbias}
\end{eqnarray}
where $W^{\alpha \beta}$ are elements of a window function matrix. Considering the first term of this expression (again, we are assuming that the linear bias term is significantly suppressed; and if this is not the case, we are simply assuming that we are setting a conservative upper limit), if $\mathbf{W}$ ends up being the identity matrix for our choices of $\mathbf{R}$ and $\mathbf{M}$, then we recover Equation \eqref{eq:super_unbiased} for the first term, and we have an estimator that has no multiplicative matrix bias. However, Equation \eqref{eq:super_unbiased} is a rather restrictive condition, and it is possible to violate it and still have a sensible (and correctly normalized) power spectrum estimate. In particular, as long as the rows of $\mathbf{W}$ sum to unity, our power spectrum will be correctly normalized. Beyond this, the data analyst has a choice for $
\textbf{M}$, and for simplicity throughout this paper we choose $\textbf{M}$ to be diagonal. In a preview of what is to come, we also stress that the derivation that leads to Equation \eqref{eq:Wpplusbias} assumes that $\mathbf{R}$ and $\mathbf{x}$ are not correlated. If this assumption is violated, a simple application of the (now incorrect) formulae in this section can result in an improperly normalized power spectrum estimator that does not conserve power, i.e., one that has signal loss.
% JEA - is one really free to choose M and R independently?

% JEA - I don't think it's attractive at all
%Without a perfect model for the covariance matrix $\textbf{C}$ of our data, one attractive way to estimate it is to empirically derive it from the data $\textbf{x}$ itself. 
Given the advantages of inverse covariance weighting, a question arises of how one goes about estimating $\C$.  One method is to empirically derive it from the data $\textbf{x}$ itself.
Similar types of weightings that are based on variance information in data are done in \citet{chang_et_al2010} and \citet{switzer_et_al2015}. In previous PAPER analyses, one time-averages the data to obtain:
\begin{equation}
\widehat{\textbf{C}} \equiv \langle\textbf{xx}^{\dagger}\rangle_{t} \approx \langle  \textbf{xx}^{\dagger}\rangle,
\end{equation}
assuming $\langle\textbf{x}\rangle_{t} = 0$ (a reasonable assumption since fringes average to zero over a sufficient 
amount of time), where $\langle \rangle_{t}$ denotes a finite average over time. The weighting matrix for our empirically estimated inverse covariance weighting is then $
\textbf{R} \equiv \widehat{\textbf{C}}^{-1}$, where we use a hat symbol to distinguish the empirical covariance from the true covariance $\textbf{C}$.

In the next three sections, we use toy models to investigate the effects of weighting matrices on signal loss by experimenting with different matrices $\textbf{R}$ and examining their impact on the resulting power spectrum estimates $\widehat{\textbf{P}}$. Our goal in experimenting with weighting is to suppress foregrounds and investigate EoR losses associated with it. We note that we purposely take a thorough and pedagogical approach to describing the toy model examples given in the next few sections. The specifics of how signal loss appears in PAPER's analysis is later described in Section \ref{sec:CaseStudy}.

%\item While the correct covariance will remove this contamination, there is no particular harm in having these modes be strongly correlated with the data, since we desire their projection in any case.

As a brief preview, we summarize our findings in the following sections here:

\begin{itemize}

\item If the covariance matrix is estimated from the data, a strong correlation between the estimated modes and the data will in general produce an estimate of the signal power spectrum which is strongly biased {\it low} relative to the true value.  In this context, this is what we call ``signal loss''  (Section \ref{sec:toymodel}).

\item The effect of the bias is worsened when the number of independent samples used to estimate the covariance matrix is reduced (Section \ref{sec:toymodel_frf}).

\item The rate at which empirical eigenvectors converge to their true forms depends on the sample variance in the empirical estimate and the shape of the empirical eigenspectrum. In general, larger sample variances lead to more loss (Section \ref{sec:toymodel_frf}).

%\item The convergence of the modes of the empirical covariance matrix to the true matrix depends on the size of the eigenvalues, with the modes with the largest eigenvalues converging more quickly (in norm) to the true eigenvector than those with smaller eigenvalues. Thus, in estimating the covariance matrix empirically, larger eigenvalue modes are likely to be more reliable (Section \ref{sec:toymodel_frf}).

\item Knowing these things, there are some simple ways of altering the empirical covariance matrix to decouple it from the data and produce unbiased power spectrum estimates (Section \ref{sec:otherweight}).

% JEA - I really struggle with whether this is intuitively obvious ... it's hard to get a sense of what "noise" is averaging down slower.  All modes have the same number of samples going into the estimate.  Should I think of the size of he eigenvalue is an estimate of its "signal/noise" in the given data realization? 
% Also, have we shown this sufficiently rigorously?  Is the toy model sufficiently general?  (Probably.)  But, for any dataset you kind of have to make a decision about where you think "reliable" is, i.e., where in eigenvalue would you make a cut?  Is Figure 20 basically trying to argue for the real data that this is ~a few modes?

\end{itemize}

\subsection{Empirical Inverse Covariance Weighting}
\label{sec:toymodel}

% JEA  - This material was moved up
%One choice for the weighting matrix $\textbf{R}$ used in power spectrum analysis is an inverse covariance matrix. This type of weighting is attractive for power spectrum analyses because it yields the smallest possible error bars on a measurement. Said differently, it gives the minimum variance estimate of the power spectrum (\citealt{tegmark_et_al1997a}; \citealt{bond_et_al1998}). In \citet{liu_tegmark2011}, it was shown that inverse covariance weighting also serves as a way to down-weight some portion of foregrounds (namely, those which do not share the same covariance structure as the cosmological signal), motivating our use of the weighting for previous PAPER analyses.
%
%One important feature of weighting data by its true inverse covariance is that, while it can suppress some foregrounds, it cannot suppress the EoR signal (nor foreground modes that masquerade as signal-like modes). By construction, inverse covariance weighting in the quadratic estimator does not lead to signal loss. Therefore, in an ideal world with perfect foreground, 
%instrumental, and EoR models, we could form $\textbf{C}$ in a way that accurately describes our measured data and use inverse covariance weighting to produce minimum variance power spectrum estimates without destroying the signal. In other words, the optimal quadratic estimator is by construction an unbiased estimator of the power spectrum.

%In practice, we do not have perfect models for $\textbf{C}$, and it is the use of an estimated covariance $\widehat{\textbf{C}}$ instead of the true covariance $\textbf{C}$ that can lead to loss. 
Using a toy model, we will now build intuition into how weighting by the inverse of the empirically estimated covariance, $\widehat{\textbf{C}}^{-1}$, can give rise to signal loss.  We construct a simple dataset that contains visibility data with $100$ time integrations and $20$ frequency channels. This model represents realistic dimensions of about an hour of PAPER data which might be used for a power spectrum analysis. For PAPER-64 (both the \citetalias{ali_et_al2015} analysis and our new analysis) we use $\sim8$ hours of data (with channel widths of $0.5$ MHz and integration times of $43$ seconds), but here we scale it down with no loss of generality.

>>>>

\documentclass[12pt,preprint]{aastex}
\usepackage[margin=1in]{geometry}
\usepackage{float,amsmath,bm,natbib,graphicx}
\citestyle{aa}

@@

\newcommand{\x}{\mathbf{x}}
\newcommand{\f}{\mathbf{f}}
\newcommand{\s}{\mathbf{s}}
\newcommand{\e}{\mathbf{e}}
\newcommand{\p}{\mathbf{p}}
\newcommand{\phat}{\widehat{\mathbf{p}}}
\newcommand{\C}{\mathbf{C}}
\newcommand{\Chat}{\mathbf{\widehat{C}}}
\newcommand{\F}{\mathbf{F}}
\newcommand{\Fhat}{\mathbf{\widehat{F}}}
\newcommand{\Q}{\mathbf{Q}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\invC}{\ensuremath{\C^{-1}}}
\newcommand{\dC}[1]{\ensuremath{\C_{,#1}}}
\DeclareMathOperator{\Tr}{tr}
\newcommand{\half}{\ensuremath{\frac{1}{2}}}
\newcommand{\PDeriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}

\begin{document}

\title{A Toy Model for Inverse Covariance Weighting} 
\author{James Aguirre, Carina Cheng, and the PAPER Collaboration}
\maketitle
\vspace{1cm}
%\section{A Toy Model for Signal Loss}

In this memo, we focus on the optimal quadratic estimator, defined as:

\begin{equation}
\label{eq:OQE}
\widehat{P}^{\alpha}  = \sum_\beta ({\F^{-1}})^{\alpha\beta} (\widehat{q}^{\beta} - \widehat{b}^{\beta} )
\end{equation}
where $\F$ is the Fisher matrix (which determines errors on the power spectrum estimate)
\begin{equation}
F^{\alpha \beta} \equiv \frac{1}{2} \textrm{tr} \left( \C^{-1} \textbf{Q}^{\alpha} \C^{-1} \textbf{Q}^{\beta} \right),
\end{equation}
$\widehat{q}$ is the un-normalized power spectrum estimate
\begin{equation}
\label{eq:OQEQuadratic}
\widehat{q}^\alpha =  \half \textbf{x}^\dagger \invC \textbf{Q}^{\alpha}  \invC \textbf{x},
\end{equation}
and $\widehat{b}$ is the additive bias
\begin{equation}
\label{eq:OQELinear}
\widehat{b}^{\alpha} = \half \Tr\left( \mathbf{U} \invC \textbf{Q}^{\alpha} \invC \right).
\end{equation}

We will mathematically illustrate its role in estimating the power spectrum of EoR. While there exists detailed literature about quadratic estimators in general (e.g., \citealt{liu_tegmark2011}; \citealt{trott_et_al2012}; \citealt{liu_et_al2014b}; \citealt{dillon_et_al2014}), here we focus on two simple cases in order to outline one situation where the estimator successfully suppresses contamination and one where it does not. By describing these two cases, we hope to clarify and motivate the desire to use OQE while also understanding its limitations.
